A Deeper Look into Bug Fixes: Patterns, Replacements, Deletions, and Additions

### 2016-02-20 MSRChallenge2016.tex
polishing

\title{A deeper look into bug fixes: Patterns, replacements, deletions, and additions}

@@

There has been considerable attention paid to automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,
  Mechtaev15,Long2016}, among many others),
%
One broad class of techniques in this space take a generate-and-validate
approach: \emph{generating} a large number of candidate patches using a
pre-defined set of mutation operators, and then \emph{validating} correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

A considerable proportion of this works targets programs written in C.
Indeed, researchers targeting Java often start from or compare against
Java-based implementations of techniques originally
implemented for C~\cite{nopol,kim2013}. % Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
%
This is an oversight because Java and C are very different languages.
Consider GenProg~\cite{legoues2012}, which
combines statement-level mutations into patches to address a particular
defect. Which ``Statements'' (a semantic unit in C) should be manipulated in Java?
Can we learn
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{Long2016}, for C)?  What types of semantic scoping
should limit mutations, given the relative stringence of the Java
compiler?

@@

\vspace{1ex}
\noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} study
a similar set of questions as we do, for C.  Beyond focusing on Java in a larger dataset~\cite{dyer2013}, we also study
replacements, ommitted from that analysis.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair, an orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique, and show that doing so results in higher-quality repairs; 
we study their templates heuristically on
a different dataset. Their results motivate
studies of human repairs, as it may result in better patches.
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new technique.  Rather, it serves as a starting
point for research on the repair of Java bugs.

@@

First, we broadly characterize bug-fixing commits to Java code.  In this paper,
we study the \emph{2015 September/Github} dataset on Boa, including $7,830,023$
Java projects with $23,229,406$ revisions. Boa identifies $4,590,679$ as bug
fixing.

@@

JAVA (ERROR) files are those that Boa failed to parse. JLS2, JLS3, and JLS4
represent Java files written in different J2SE versions.  We see that text and
Binary files are changed least frequently. This is unsurprising, since text
files are often documentation, and binaries should be changed rarely. If we
consider all 4 kinds of Java files as one kind, on average, each bug fixing
revision changes $4.47$ Java files. This does not conform to our assumption of
bug locality. XML files in Java projects usually represent build files;
changing build files is unsurprising.  Rather more surprising is how frequently
UNKNOWN files are changed. Given that all projects are Java projects, it is
unclear whether these are written in other programming languages. We
hypothesize they might be test resources.  To better understand what changes
are usually made to Java files, we next ask:

@@

Boa's capabilities are powerful, but limited in the precision it enables in
our detection of Par bug-fixing patterns. For example, it cannot directly \texttt{diff}
two files versions. Thus, rather than finding the exact count of
bug fixing patterns, we approximate by processing pre- and
post- fix files separately.  Additionally, there are
two of the 10 patterns that cannot be easily detected by
Boa, as we describe below. 
%
The patterns we search for are as follows, including our mechanism for
detecting them:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This template changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre-fix and post-fix versions
of a buggy file, we create a custom method call signature that
contains the name, literal parameters, and variable parameters (complex
expressions are listed as OTHER).
%  This is due to inexistence of
% function that can print Abstract Syntax Tree (AST) back to source code in
% Boa. 
We discard method signatures that appear both in pre-fix and post-fix
versions, and then identify methods with the same name and the
same number of parameters, but different signatures between versions.

@@

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
template adds a parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create a method signature similar to the above, seeking methods with
  the same name, one more parameter in the post-fix version.

\vspace{1ex}
\noindent
 {\bf Changing a branch condition (CBC)}. This template adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
We count the number of { \em logical and} and {\em
    logical or} inside if conditional expression sfor both pre-fix and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a larger count in post-fix version,
  we consider the CBC pattern is found.

\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This template adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that IAO pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This template inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre-fix and post-fix
  version. If there is a larger count in post-fix version,
  we consider that ANC pattern found.

\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
%	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
 We count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre-fix and
  post-fix version. If there is a larger count in post-fix version, we consider that AOB pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre-fix and
post-fix version. If the count is different between pre- and post-fix
version, we consider the COB pattern found.

\vspace{1ex} The above patterns cover 8 out of 10 bug fixing patterns in
PAR. The other two patterns are (1) Class Cast Checker (cannot be analyzed because
Boa does not support the {\em instanceof} expression); and (2) Expression Changer (requires us to track scope between pre-fix and post-fix version; not
easily supported by Boa). Investigating 8 out of
10 patterns still shows us how common the patterns actually are.

@@

\begin{table}[!htb]
	\centering
	\begin{tabular}{lrr} 
		\hline
		& \textbf{\#Appearance} & \textbf{Percentage}\\
		\hline
		%Buggy Files & 46,301,429 \\ 
		AMP Pattern & 901,083 & 1.95\%\\
		MSM Pattern & 783,073 & 1.69\%\\
		COM Pattern & 270,128 & 0.58\%\\ 
		CBC Pattern & 1,959,377 & 4.23\%\\  
		IAO Pattern & 1,308,006 & 2.82\%\\  
		ANC Pattern & 1,340,561 & 2.90\%\\  
		AOB Pattern & 128,016 & 0.28\%\\  
		COB Pattern & 262,915 & 0.57\%\\   
		\hline
	\end{tabular}
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}

@@

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is CBC (changing a branch condition); and the least common
pattern is AOB (adding an array out of bound checker). If we conservatively assume that these patterns never appear together, they
cover 14.78\% of buggy files. In the dataset studied by Kim et al., the overall
10 patterns cover almost 30\% of real patches~\cite{kim2013}. Although we only analyze 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and likely represents an upper
bound.  This suggests that these fixing patterns might not generalize
beyond an initial dataset, and that more work remains to expressively characterize human bug-fixing behavior.

@@

\begin{figure*}[!t] \centering \includegraphics[width=\textwidth]{StmtType} \caption{For each kind of statement, how many
	bug fixing revisions add/delete statements of the same kind?}
	\label{fig:StmtType} \end{figure*}

@@

\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the DoStatement row (row 5) and AssertStatement column (column 2)
shows 0.81, meaning that Do statements were
replaced by Assert statements 0.81\% of the times.

@@

\noindent\textbf{Threats to validity.} Since we are using a DSL
provided by Boa, the correctness of our analysis
depends on both our programs and Boa itself. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem.  To mitigate the risk of errors in
implementation, we open source our scripts~\footnote{\url{https://github.com/chupanw/BoaChallenge}}.
Because Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions, our template
matching and analysis of code changes (by counting each statement kind) only
provide estimates of behavior; we consider our results as informative
approximations.

>>>>

\title{A Deeper Look into Bug Fixes: Patterns, Replacements, Deletions, and Additions}

@@

There has been considerable attention paid to automatic
program repair in recent
years~\cite{kim2013,legoues2012,pan2009,
  Mechtaev15,Long2016}.
%
One broad class of techniques in this space take a generate-and-validate
approach: \emph{generating} a large number of candidate patches using a
pre-defined set of mutation operators, and then \emph{validating} correctness with
respect to a set of test cases.  Mutation operators range from simple, coarse
granularity statement-level mutations to human-constructed templates learned
from a large corpus of previous human bug-fixing commits.

A considerable proportion of these works target programs written in C.
Indeed, researchers targeting Java often start from or compare against
Java-based implementations of techniques originally
implemented for C~\cite{nopol,kim2013}. % Even PAR~\cite{kim2013}, perhaps the
% best-known previously-proposed Java-based program repair technique, both starts
% from and is directly compared to a Java implementation of GenProg~\cite{genprog}, which
% targets the C programming language.  
%
This is an oversight because Java and C are very different languages.
Consider GenProg~\cite{legoues2012}, which
combines statement-level mutations into patches to address a particular
defect. Which ``Statements'' (a semantic unit in C) should be manipulated in Java?
Can we learn
intelligent rules to inform candidate modifications
(cf. Prophet~\cite{Long2016}, for C)?  What types of semantic scoping
should limit mutations, given the relative stringence of the Java
compiler?

@@

\vspace{1ex}
\noindent\textbf{Related work.} Zhong and Su~\cite{zhong2015} study
a similar set of questions as we do, for C.  Beyond focusing on Java in a larger dataset~\cite{dyer2013}, we also study
replacements, omitted from that analysis.  Barr et al.~\cite{Barr14fse} studied changes to
Java programs to understand the ``Plastic Surgery Hypothesis'' underlying
certain types of program repair, an orthogonal
concern. Kim et al.~\cite{kim2013} manually analyze 
changes to Java programs to inform an automated repair
technique, and show that doing so results in higher-quality repairs; 
we study their templates heuristically on
a different dataset. Their results motivate
studies of human repairs, as it may result in better patches.
Long and Rinard learn probabilistic models from bug fixes to C
code~\cite{Long2016}. Our analysis is significantly less precise and does not
inform a new technique.  Rather, it serves as a starting
point for research on the repair of Java bugs.

@@

First, we broadly characterize bug-fixing commits to Java code.  In this paper,
we study the \emph{September 2015/Github} dataset on Boa, including $7,830,023$
Java projects with $23,229,406$ revisions. Boa identifies $4,590,679$ as bug
fixing.

@@

JAVA (ERROR) files are those that Boa failed to parse. JLS2, JLS3, and JLS4
represent Java files written in different J2SE versions.  We see that text and
binary files are changed least frequently. This is unsurprising, since text
files are often documentation, and binaries should be changed rarely. If we
consider all 4 kinds of Java files as one kind, on average, each bug fixing
revision changes $4.47$ Java files. This does not conform to our assumption of
bug locality. XML files in Java projects usually represent build files;
changing build files is unsurprising.  Rather more surprising is how frequently
UNKNOWN files are changed. Given that all projects are Java projects, it is
unclear whether these are written in other programming languages. We
hypothesize they might be test resources.  To better understand what changes
are usually made to Java files, we next ask:

@@

Boa's capabilities are powerful, but limited in the precision it enables in
our detection of PAR bug-fixing patterns. For example, it cannot directly \texttt{diff}
two files versions. Thus, rather than finding the exact count of
bug fixing patterns, we approximate by processing pre- and
post-fix files separately.  Additionally, there are
two of the 10 patterns that cannot be easily detected by
Boa, as we describe below. 
%
The patterns we search for are as follows, including our mechanism for
detecting them:

\vspace{1ex}
\noindent {\bf Altering method parameters (AMP)}.  This template changes 
input method parameters. %(\texttt{obj.method(v1,v2) $\rightarrow$
%obj.method(v1,v3)}). 
To detect this pattern, for both pre- and post-fix versions
of a buggy file, we create a custom method call signature that
contains the name string, literal parameters string, and variable parameters string (complex
expressions are listed as ``OTHER'' string).
%  This is due to inexistence of
% function that can print Abstract Syntax Tree (AST) back to source code in
% Boa. 
We discard method signatures that appear both in pre- and post-fix
versions, and then identify methods with the same name and the
same number of parameters, but different signatures between versions.

@@

\vspace{1ex}
\noindent
{\bf Calling another overloaded method with one more parameter (COM)}. This
template adds a parameter to a method. %(\texttt{obj.method(v1) $\rightarrow$ obj.method(v1,v2)}).
We create a method signature similar to the above, seeking methods with
  the same name, but with one more parameter in the post-fix version.

\vspace{1ex}
\noindent
 {\bf Add or remove a branch condition (ABC)}. This template adds or removes a condition.
%(e.g., \texttt{if(a != b) $\rightarrow$ if(a != b \&\& c == 0)}).
We count the number of { \em logical and} and {\em
    logical or} inside if conditional expression for both pre- and post-fix
  version. We assume that the addition/removal of logical operators indicate
  addition/removal of condition. If there is a count difference between pre- and post-fix version,
  we consider an instance of the ABC pattern is found.

\vspace{1ex}
\noindent
{\bf Initializing an object (IAO)}. This template adds a initialization to object
declaration.  We count the number of NEW expressions in variable declarations
for both pre- and post-fix versions. If there is a larger count in post-fix
version, we consider that an instance of the IAO pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a null checker (ANC)}.  This template inserts a condition to check whether an
    object is null.
%	{\bf Example:} obj.m1() $\rightarrow$ if(obj!=null){obj.m1()}
  To detect ANC pattern, we count the number of if conditional expressions that
  contains {\em !=null} or {\em ==null} for both pre- and post-fix
  version. If there is a larger count in post-fix version,
  we consider that an instance of the ANC pattern found.

\vspace{1ex}
\noindent
 {\bf Adding an array out of bound checker (AOB)}. This pattern inserts a condition to check that an array
    index is within bound right before the index is used.
%	{\bf Example:} arr[idx]=0 $\rightarrow$ if(idx<arr.length){arr[idx]=0}.
 We count the number of if conditional expression that
  contains {\em expr<var.length} or {\em var.length>expr} for both pre- and
  post-fix version. If there is a larger count in post-fix version, we consider that an instance of the AOB pattern is found.

\vspace{1ex}
\noindent
 {\bf Adding a collection out of bound checker (COB)}.  This pattern inserts a condition to check that a
    collection index is within bound before it is used.
To detect COB pattern, we count the number of if conditional expression that
contains {\em expr<col.size()} or {\em col.size()>expr} for both pre- and
post-fix version. If the is a larger count in post-fix
version, we consider that an instance of the COB pattern is found.

\vspace{1ex} The above patterns cover 8 out of 10 bug fixing patterns in
PAR. The other two patterns are (1) Class Cast Checker (cannot be analyzed because
Boa does not support the {\em instanceof} expression); and (2) Expression Changer (requires us to track scope between pre- and post-fix version; not
easily supported by Boa). Investigating 8 out of
10 patterns still shows us how common the patterns actually are.

@@

\begin{table}[!htb]
	\centering
	\begin{tabular}{lrr} 
		\hline
		& \textbf{\#Appearance} & \textbf{Percentage}\\
		\hline
		%Buggy Files & 46,301,429 \\ 
		AMP Pattern & 901,083 & 1.95\%\\
		MSM Pattern & 783,073 & 1.69\%\\
		COM Pattern & 270,128 & 0.58\%\\ 
		ABC Pattern & 1,959,377 & 4.23\%\\  
		IAO Pattern & 1,308,006 & 2.82\%\\  
		ANC Pattern & 1,340,561 & 2.90\%\\  
		AOB Pattern & 128,016 & 0.28\%\\  
		COB Pattern & 262,915 & 0.57\%\\   
		\hline
	\end{tabular}
	\caption{Frequency of Bug Fixing Patterns}\label{tab:freqpattern}

@@

Table~\ref{tab:freqpattern} estimates the prevelance of PAR templates in the
Java dataset. The most common pattern we observed is ABC (add or remove a branch condition); and the least common
pattern is AOB (adding an array out of bound checker). If we conservatively assume that these patterns never appear together, they
cover 14.78\% of buggy files. In the dataset studied by Kim et al., the overall
10 patterns cover almost 30\% of real patches~\cite{kim2013}. Although we only analyze 8
out of 10 bug fixing patterns, it is unlikely the final two will cover the
difference. Moreover, our detection is imprecise, and might represent an upper
bound.  This suggests that these fixing patterns might not generalize
beyond an initial dataset, and that more work remains to expressively characterize human bug-fixing behavior.

@@

\begin{figure*}[!t] \centering \includegraphics[width=\textwidth]{StmtType}
	\vspace{-0.7cm} \caption{For each kind of statement, how many
	bug fixing revisions add/delete statements of the same kind?}
	\label{fig:StmtType} \end{figure*}

@@

\noindent\textbf{Potential replacements.} Table~\ref{tab:likeliness} shows
the replacement likelihood for our dataset 
(each cell shows the percent of the time that the statement in the \emph{row}
was replaced by a statement of the type in the \emph{column}). For example, if
the cell in the DoStatement row (row 5) and AssertStatement column (column 2)
shows 0.81, it means that Do statements were
replaced by Assert statements 0.81\% of the times.

@@

\noindent\textbf{Threats to validity.} Since we are using a DSL
provided by Boa, the correctness of our analysis
depends on both our programs and Boa itself. For example, we rely on Boa to
identify bug fixing revisions; however, precisely accomplishing this is an open
problem.  To mitigate the risk of errors in
implementation, we open source our scripts\footnote{\url{https://github.com/chupanw/BoaChallenge}}.
Because Boa does not provide an easy mechanism
to identify precise, statement-level diffs between revisions, our template
matching and analysis of code changes (by counting each statement kind) only
provide estimates of behavior; we consider our results as informative
approximations.
